Title

### 2013-04-01 inc/model_results/shapes/shapes.tex
clear some TODOs from shapes
tweaks to shapes

For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects can dominate the propagation dynamics.
\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
%TODO revisit this next sentance, the data doesn't seem to bear it out
In the spherical case, space-charge induced bunch broadening begins to overwhelm the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing a more uniform three-dimensional `Coulomb explosion' of the electron pulse with the overall bunch broadening rate much greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation. This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
%TODO the following sentance appears not to be true, still perhaps should include something like it
%In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.

The data presented in \ref{fig:compare_shape} using the AG model of Michalik and Sipe \cite{michalik_analytic_2006} are entirely consistent with prior works.\cite{reed_femtosecond_2006,siwick_ultrafast_2002}
In femtosecond UED experiments, significant efforts are made to offset space-charge induced temporal (longitudinal, $z$) pulse broadening by reducing the time of flight from the gun photocathode to the sample.\cite{siwick_ultrafast_2002,reed_evolution_2009}
Moreover, efforts are now underway to compensate for temporal electron pulse broadening using RF cavity pulse compression techniques.\cite{oudheusden_electron_2007}
In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers,\cite{lobastov_four-dimensional_2005} deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
Even for nanosecond DTEM (the extreme prolate pulse case, $\xi \gg 1 $), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
In this thesis, however, I will restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}

>>>>

For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects can dominate the propagation dynamics.
\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
In the spherical case, space-charge induced bunch broadening begins to overwhelm the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing a more uniform three-dimensional `Coulomb explosion' of the electron pulse with the overall bunch broadening rate much greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation. This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
% the following sentance is no longer true, still perhaps should include something like it
%In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.

The data presented in \ref{fig:compare_shape} using the AG model of Michalik and Sipe \cite{michalik_analytic_2006} are entirely consistent with prior works.\cite{reed_femtosecond_2006,siwick_ultrafast_2002}
In femtosecond UED experiments, significant efforts are made to offset space-charge induced temporal (longitudinal, $z$) pulse broadening by reducing the time of flight from the gun photocathode to the sample.\cite{siwick_ultrafast_2002,reed_evolution_2009}
Moreover, efforts are now underway to compensate for temporal electron pulse broadening using RF cavity pulse compression techniques.\cite{oudheusden_electron_2007}
In UEM, employing a standard retro-fitted electron microscope column driven by MHz repetition-rate sub-picosecond lasers,\cite{lobastov_four-dimensional_2005} deleterious space-charge effects have obliged operation in the $\sim$1 electron/pulse regime to maintain high spatial resolution.
Even for nanosecond DTEM (the extreme prolate pulse case, $\xi \gg 1 $), transverse space-charge induced beam broadening has necessitated the insertion of an additional magnetic electron lens in a microscope column to allow more efficient delivery of electrons to the specimen when $ N > 10^{6} $.\cite{lagrange_nanosecond_2008}
In this thesis, however, I will often restrict the discussion to ultrashort (ps and sub-ps) electron pulses which (i) allow for the temporal resolution of fundamental events occurring on ultrafast time scales in materials science, biology and chemistry \cite{king_ultrafast_2005} and (ii) can be easily generated in the oblate regime ($ \xi < 1 $) where transverse bunch broadening effects are minimized (\ref{fig:compare_shape})(enabling more efficient beam propagation) and RF cavities can be used to compensate and reverse longitudinal (temporal) pulse broadening.\cite{veisz_hybrid_2007}
